\title{Thursday Warm-Up (INTD262)}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[12pt]{article}
\usepackage[margin=1.25cm]{geometry}
\usepackage{hyperref}
\begin{document}
\maketitle

\section{Chapter 5 of \textit{The Scientific Attitude}}

\begin{enumerate}
\item Define the term \textit{confirmation bias}. \\ \vspace{0.5cm}
\item \textbf{Recall the story of cold fusion} in Chapter 5.  (a) List three facets of the peer review process that went wrong in this episode. \\ \vspace{2cm}
\item How long, from start to finish, did it take for the scientific community to sort out the errors in the cold fusion research? \\ \vspace{1cm}
\item Consider the passages in chapter 5 about scientists withholding raw data when publishing. \\
(a) Make a determination about the logic of the following statement:
\begin{quote}
Fraudulent results are almost always irreproducible.  Therefore, irreproducible studies are almost always indicate fraud.
\end{quote}
(b) What fraction of psychological studies were found to be irreproducible?\footnote{Ian Sample, ``Study Delivers Bleak Verdict on Validity of Psychology Experiment Results,'' \textit{Guardian}, Aug. 27th, 2015.} \\ \vspace{1cm}
\item Did the conclusion drawn from the paper in the previous question pan out?  Why or why not? \\ \vspace{1cm}
\item If you set out to reproduce the results of an experiment, are you guaranteed to reproduce it if you follow the methods of the study exactly?
\end{enumerate}

\end{document}
