\title{Tuesday Warm-Up (INTD262)}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[12pt]{article}
\usepackage[margin=1.25cm]{geometry}
\usepackage{hyperref}
\begin{document}
\maketitle
\small

\section{Chapter 4 of \textit{Science in Latin America}}

\begin{enumerate}
\item Note that we encountered several examples of viceregal engineers becoming Latin American leaders in Chapter 4.  (a) What are some examples of professions that involved modern technical skill in R\'{i}o de la Plata and Per\'{u}? (b) What is the primary profession of modern American leaders, for example, elected to Congress? \\ \vspace{1.5cm}
\item Jos\'{e} Mariano Moci\~{n}o and others were ordered by the Mexican viceroy on an expedition to Nootka Island.  What was the purpose of the expedition?  Were you aware that scientists from Nueva Espa\~{n}a ventured this far North? (Take INTD255 to learn more!) \\ \vspace{1cm}
\item In Per\'{u}, we must take note of the work of Hip\'{o}lito Un\'{a}nue.  For example, he was one of the founders of \textit{La Sociedad Acad\'{e}mica de Amantes del Pa\'{i}s} in 1790.  (a) What are some of his other scientific contributions? (b) In Nueva Granada, we must take note of the work of Jos\'{e} Celestino Mutis.  What are some of his main contributions? \\ \vspace{1.25cm}
\item In Chapter 4, we encounter the following quote:
\begin{quote}
The discovery of America had extraordinary scientific and cultural consequences.  The firm belief that there remained no territories to be discovered made the world less mysterious and no longer a boundless space; it came to be seen as complete and therefore susceptible to order and measurement.
\end{quote}
Reflect: \textit{Who} discovered America, and how susceptible is the world to order and measurement? \\ \vspace{0.5cm}
\item In Chapter 4, we encounter the Californian product \textit{cochineal}.  Do you know what it is?
\item The authors note the effect that Latin American science had on ``the home country,'' or ``peninsular'' people.  Can you give an example?
\end{enumerate}

\end{document}
