\title{Reading Quiz 0.1 - INTD290}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[10pt]{article}
\usepackage[a4paper, total={18cm, 27cm}]{geometry}
\usepackage{outlines}
\usepackage{hyperref}
\usepackage{graphicx}
\begin{document}
\maketitle

\section{How to Submit this Assignment}

Once you answer the questions, take a picture of your work and convert it to a PDF.  Submit the PDF to the assignment link on Moodle.

\section{Introduction - The Latin American Scientific Theater (Juan Jos\'{e} Salda\~{n}a)}

\begin{enumerate}
\item The author begins by describing different methods of historical analysis as applied to Latin America, and in particular scientific development there.  What is the author's conclusion about the social history of Latin America with respect to science? \\ \vspace{2cm}
\item What is an example of the failure of economic explanations for the development of science given by the author? \\ \vspace{2cm}
\item What is the function of La Sociedad Latinoamericana de Historia de las Ciencias y Tecnolog\'{i}a, when was it founded, and what does it publish? \\ \vspace{1cm}
\item What does the concept of \textit{peripheral science}, versus \textit{central science} have to do with Latin America and Europe?  What are some examples the author gives in which Latin America was \textit{central}? \\ \vspace{2cm}
\end{enumerate}

\section{Chapter 1 - Natural History and Herbal Medicine in Sixteenth-century America (Xavier Lozoya), part I}

\begin{enumerate}
\item What does the author have to say about the two versions of the history of scientific thought in Latin America, with regard to the Enlightenment? When did Christopher Columbus land in the New World, and what was the origin of the word ``Indian?'' In the opinion of the author, what will be the future of indigenous history? \\ \vspace{2cm}
\item From the viewpoint of the author, what seems to be the definition of the term ``ethnohistory?''  How is it connected to globalization? \\ \vspace{2cm}
\item How was the reconquest of former Spanish territory held by Arabs from North Africa different from the conquest of Latin America with regard to preservation or censorship of knowledge? \\ \vspace{1cm}
\item Describe the origin of the myth of the Fountain of Youth.  How did this tale arise in both Europe and Latin America? What does it tell you about \textit{how people interpret what they see and hear in a foreign land}? \\ \vspace{1cm}
\item Upon hearing of the condors from indigenous people, European colonials thought the description matched which animal?  Why?
\end{enumerate}

\end{document}
