\title{Thursday Warm-Up (INTD262)}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[12pt]{article}
\usepackage[margin=1.25cm]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{subcaption}
\begin{document}
\maketitle

\small

\section{M. P. Ramos-Lara: Contribuciones de astr\'{o}nomos mexicanos al estudio de auroras boreales de baja latitud entre 1789 y 1791}

\begin{enumerate}
\item What did Jos\'{e} Antonio Alzate y Ram\'{ı}rez do when he and a friend first saw the aurora in 1789? \\ \vspace{0.75cm}
\end{enumerate}

\section{Chapter 4 of \textit{Science in Latin America}}

\begin{enumerate}
\item (a) Name three organizations that sponsored scientific expeditions throughout Latin America in the Enlightenment Period.  (b) What does the author mean by the following quote: ``And how do we explain the high percentage of reformers dressed in cassocks in Hispanic America if we continue to insist on the contradiction between the Enlightenment and religion?'' \\ \vspace{1.25cm}
\item What was notable about the writing of the Benedictine monk Benito Jer\'{o}nimo Feijoo? \\ \vspace{1cm}
\item What was notable about the explorations of Jos\'{e} Sanchez Labrador? \\ \vspace{1cm}
\item Why do you suppose there is so much discussion of the Jesuit cartography and natural history surrounding rivers in R\'{i}o de la Plata and Per\'{u}? \\ \vspace{1cm}
\item (a) List several scientific or engineering functions of the viceroyal officials in Enlightenment Latin America.  How did they disseminate their ideas for educational reform? (b) As the term ``engineer'' became more widely used, what examples of infrastructure were built by this new class of worker? \\ \vspace{1cm}
\end{enumerate}

\end{document}
