\title{Reading Quiz 0.2 - INTD290 ... Part I}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[10pt]{article}
\usepackage[a4paper, total={18cm, 27cm}]{geometry}
\usepackage{outlines}
\usepackage{hyperref}
\usepackage{graphicx}
\begin{document}
\maketitle

\section{How to Submit this Assignment}

Once you answer the questions, take a picture of your work and convert it to a PDF.  Submit the PDF to the assignment link on Moodle.

\section{Hispanic America's Historical Dynamics}

\begin{enumerate}
\item The author lays out the central argument of the chapter in this section, which is the connection between merchant guilds, the domestic economy, and the constitution of enlightenment institutions.  Please list some of the institutions, like colleges and botanical gardens (think of these as research institutes), mining guilds and corporations, etc.  Give the dates, organized by country below:
\begin{enumerate}
\item M\'{e}xico: 
\vspace{1cm}
\item Per\'{u}:
\vspace{1cm}
\item Venazuela (Caracas):
\vspace{1cm}
\item Guatemala:
\vspace{1cm}
\end{enumerate}
\end{enumerate}

\section{Modern Scientific Culture}

\begin{enumerate}
\item Discuss the significance of \textit{empiricism}, or the use of observations from instruments, on scientific culture.  How was this different from Scholasticism? (Feel free to look up the definitions of these terms). \\ \vspace{2cm}
\end{enumerate}

\section{Hispanic American Scientific Culture}

\begin{enumerate}
\item Notice the discussion of \textit{Scientific Libraries}.  Discuss the significance of the formation of private scientific libraries in Enlightenment period Hispanic America.  What does the existence of these collections of new, modern books tell us about scientific culture at the time?  Why where these books not in common libraries?  What are a few other examples of the birth of modern scientific culture that more private than public or nationalized? \vspace{3cm}
\item Notice the discussion of \textit{Scientific and Technical Journals}.  Give several examples of scientific and technical journals emerging in the culture.  When were they created and what did they cover? \vspace{2cm}
\item What was the significance of the debate over mining methods?  Who eventually won the debate and why?
\end{enumerate}

\end{document}
