\title{Constitutional Comparison in the Context of Scientific and Educational Progress}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[12pt]{article}
\usepackage[margin=1.25cm]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\begin{document}
\maketitle
\small

\section{Introduction}

In the following activity, we will locate provisions related to scientific progress and education within the Mexican and US Constitutions and subsequent laws.  The goal is to compare and contrast the legal and scientific styles present in both.

\section{The Mexican Foundational Documents}

Recall what we have read in chapter 5 of \textit{Science in Latin America} about the early foundational documents of the Mexican nation-state.  Here are links to several of them:
\footnotesize
\begin{itemize}
\item The 1814 Apatzing\'{a}n Constitution (I struggled to find this one in English).  References in \textit{Science in Latin America}.
\item The 1824 Mexican Constitution: \\ \url{https://tarlton.law.utexas.edu/constitutions/federal-constitution-united-mexican-states-1824}.
\item The 1917 Mexican Constitution \\ \url{https://www.constituteproject.org/constitution/Mexico_2015}.
\end{itemize} \vspace{1cm}
\small
\section{The US Foundational Documents}

Recall your basic knowledge of the early foundational documents of the United States.  Here are links to the Declaration of Independence and the US Constitution of 1787:
\footnotesize
\begin{itemize}
\item The Declaration of Independence: \url{https://www.archives.gov/founding-docs/declaration-transcript}.
\item The United States Constitution of 1787: \url{https://www.archives.gov/founding-docs/constitution-transcript}.
\end{itemize} \vspace{1cm}
\small
\textbf{Our task is simple.} Find all references to science, technology, engineering, mathematics, and education in these documents.  Compare and contrast the context and purpose of these passages, and discuss them with your classmates.

\end{document}
